---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
# Load necessary libraries
library(dplyr)    # For data manipulation
library(ggplot2)  # For plotting
library(knitr)    # For knitting
library(readr)    # If you're reading data from CSV files or similar formats

# Set knitr options for chunk output
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ufoPackage

<!-- badges: start -->
<!-- badges: end -->

The ufoPackage aims to provide tools for exploring and analyzing UFO sighting data. This package includes a cleaned dataset, a Shiny app for interactive data exploration, and a set of functions designed to facilitate further analysis.

## Installation

You can install the development version of ufoPackage from [GitHub](https://github.com/rows0001) with:

```{r}
devtools::install_github("ETC5523-2024/assignment-4-packages-and-shiny-apps-rows0001/ufopackage/ufoPackage")
```


## Data

The package includes a dataset containing UFO sightings, featuring various attributes such as reported date, location, shape, and time of sighting. The data is cleaned and processed to ensure consistency and usability.



## Usage

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ufoPackage)
## basic example code
```


## Load Data

Load the cleaned UFO sightings data:

```{r}
data("ufo_sightings")  # Load the UFO sightings dataset

```


## Summary of Sightings

You can get a summary of the sightings data:

```{r}
summary(ufo_sightings)

```



## Plotting Sightings
To visualize sightings over the years:

```{r}
# Summarize sightings per year
yearly_sightings <- ufo_sightings %>%
  mutate(year = format(as.POSIXct(reported_date_time, format = "%Y-%m-%d %H:%M:%S"), "%Y")) %>%
  group_by(year) %>%
  summarise(count = n(), .groups = 'drop')

# Plot
ggplot(yearly_sightings, aes(x = as.numeric(year), y = count)) +
  geom_line() +
  labs(title = "UFO Sightings per Year", x = "Year", y = "Number of Sightings")

```


## Shiny App

This package includes a Shiny app that allows users to interactively explore the UFO sightings dataset. You can launch the app using the following command:

```{r}
#run_ufo_app()  # This function launches the Shiny app

```


## Shiny App Features

- Interactive Selection: Users can filter sightings based on different criteria such as year, shape, and location.
- Dynamic Outputs: The app updates visualizations and summaries based on user selections.
- User Guidance: Clear descriptions are provided to help users understand the dataset and how to interpret the outputs.


## pkgdown Site

For more detailed documentation, visit the [pkgdown site](https://etc5523-2024.github.io/assignment-4-packages-and-shiny-apps-rows0001/).



## Contribution

If you would like to contribute to ufoPackage, feel free to fork the repository and submit a pull request. For any issues or feature requests, please create an issue in the GitHub repository.


## License

This project is licensed under the MIT License - see the LICENSE file for details.

## Acknowledgements

Special thanks to the sources of the UFO dataset and contributors to the R community for their valuable resources.

